P.berghei pir analysis

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
added more functions to separate files

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
```

```{r function_mround}
#Make nicer numbers rounded up, useful for making scales for a plot.
mround <- function(x,base){ 
  base*ceiling(x/base) 
} 
```

```{r function_my_max}
#max robust to NAs
my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) 
```

```{r function_firstup}
#Capitalize first letter
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}
```

@@

```{r transcript_redundancy, message = FALSE, results='hide'}

#Note that there is some redundancy
two.transcript <- allgene.info %>% 
  dplyr::count(Geneid) %>% 
  filter(n > 1)

#Nine genes have two transcripts - need to give them one average of the transcripts since featurecounts has only
# calculated one count for the locus
two.transcript.info <- allgene.info[allgene.info$Geneid %in% two.transcript$Geneid,]
two.transcript.mean = aggregate(transcript_length~Geneid, data=two.transcript.info, 
                                FUN=function(x) c(mean=mean(x)))

#Remove the first match
allgene.info <- allgene.info[-(match(two.transcript.mean$Geneid, 
                                     allgene.info$Geneid)),]

#For the second match replace the transcript value with the mean calculated with aggregate
allgene.info$transcript_length <- as.double(allgene.info$transcript_length)
allgene.info[match(two.transcript.mean$Geneid, 
                   allgene.info$Geneid),]$transcript_length <- two.transcript.mean$transcript_length
lengths.transcript <- subset(allgene.info, select = c(Geneid, transcript_length), drop = F)
```

## Read in counts and calculate TPM

```{r read_files_calc_TPM}
#Reading in the counts from file
berghei_counts <- read.csv('tl20-05_berghei_counts.csv',
                               row.names = 1)

berghei_tpm <- tpm(berghei_counts,
                   lengths = lengths.transcript)

apply(berghei_tpm[-1], 2, sum)
```

```{r save_tpm_oldNames_allSamples}
#Save the 'raw' TPM before any samples have been removed or sample names combined.

write.csv(berghei_tpm,
          'berghei_tpm_oldNamesAllSamples.csv')

```{r remove_E06-6h}
#Remove the Caldelari (E06) 6h samples, as they did in their paper.
berghei_tpm[,str_subset(colnames(berghei_tpm), 'E06_Liv.6h')] <- NULL
```

```{r hclusts, message = FALSE}
experiment_id <- unique(na.omit(str_extract(colnames(berghei_tpm), pattern = 'E[[:digit:]]{2}')))

lapply(experiment_id, 
       function(experiment_name){
         tpm_data <- select(berghei_tpm, matches(experiment_name))
         if(ncol(tpm_data) > 3){
           dists = dist(t(tpm_data))
           par(cex=0.7)
           plot(hclust(dists),
                main = experiment_name)
         }
       })

Going to combine certain samples based on their similar profiles, e.g. <4h, >48h liver stages.

```{r combine_similar_samples_byNames}
col_stage <- colnames(berghei_tpm)

#Combine all Liver samples 4h or before into LiverPre4h
col_stage <- str_replace_all(col_stage,
                             pattern = 'Liv\\.(2|4)h_[[:digit:]]+',
                             replacement = 'Liv.pre4h')
#Combine all Liver samples 48h or onwards into LiverPost48h
col_stage <- str_replace_all(col_stage,
                             pattern = 'Liv\\.(4[8-9]|[5-9][0-9])h_[[:digit:]]+',
                             replacement = 'Liv.post48h')
#Combine all blood meal samples 3h or before into BldMlPre3h
col_stage <- str_replace_all(col_stage,
                             pattern = '(1h|3h)_[[:digit:]]+', 
                             replacement = 'Pre3h')
#Uniquify the name names
col_stage[grep(col_stage, 
               perl = TRUE,
               invert = TRUE,
               pattern = '_[[:digit:]]{1,}$')] <- uniquify(col_stage[grep(col_stage, 
                                                                          perl = TRUE,
                                                                          invert = TRUE,
                                                                          pattern = '_[[:digit:]]{1,}$')]) 
```

```{r remove_badly_clustering_samples}
berghei_tpm$`E08_Liv.pre4h_1` <- NULL
berghei_tpm$`E03_Asex.Schiz_1` <- NULL
berghei_tpm$`E03_Asex.Schiz_3` <- NULL
```

@@

```{r save_tpm}
write.csv(berghei_tpm,
          'berghei_tpm.csv')
```

@@

lapply(all_top, `[`, c('all_genes', 'pir_genes'))

>>>>

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
```

@@

@@

```{r save_tpm}
berghei_tpm <- read.csv('berghei_tpm.csv', 
                        row.names = 1)
```

@@

lapply(all_top[[1]], `[`, c('all_genes', 'pir_genes'))
